---
title: "Set01Answer"
author: "Yaoyao"
date: "January 27, 2015"
output: pdf_document
---

```{r}
library(foreign)
set1<-read.csv("~/Documents/Data/Set01.csv")
names(set1)
x11()
split.screen(c(1,2))
screen(1)
boxplot(set1$y,main="y")
screen(2)
boxplot(set1$x1,set1$x2,set1$x3,main="x1 x2 x3")
```

y and x1 x2 x3 are in different scale. Y has a large outlier observation. So I delete the outlier first.

``` {r}
set1$y[set1$y > quantile(set1$y,probs=c(.25,.75))[2]+1.5*IQR(set1$y)]<-NA
```

I then plot descriptive scatter plots: 
```{r}
x11()
split.screen(c(2,2))
screen(1)
plot(set1$y~set1$x1,pch=20)
screen(2)
plot(set1$y~set1$x2,pch=20)
screen(3)
plot(set1$y~set1$x3,pch=20)
```

It seems x1 and y are positively correlated, x2 and y are negatively correlated. And the relationship between x3 and y is unclear. I then did a simple linear regression: 
```{r}
model<-lm(set1$y~set1$x1+set1$x2+set1$x3)
summary(model)
```

The results confirm what we saw in the scatter plots.All three independent variables are significantly correlated with y. 


